---
title: "factor-xmaps"
author: "cynthiahqy"
date: "2023-01-18"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Background

-   Factors are the base R S3 class for categorical variables
-   Factors are integer vectors with a levels attribute.
-   forcats provides convenience and safe-guarded functions for working with factors

## Visualisation Idea

-   Take levels before and after transformation, generate and visualise the implicit `xmap`.

```{r}
fct <- factor(c("a", "b", "b", "c", "d"))

after_other <- fct |>
  forcats::fct_other(keep = c("a", "b"), other_level = "other")
stopifnot(identical(after_other, factor(c("a", "b", "b", "other", "other"))))
```

For reference this is the mapping:

```{r}
xmap_ref <-
  tibble::tribble(~key1, ~key2, ~weight,
                  "a", "a", 1,
                  "b", "b", 1,
                  "c", "other", 1,
                  "d", "other", 1)
```

We can probably only extract the mapping if there is at most one target node involved in a many-to-one mapping.

```{r}
collapse_xmap <- tibble::tibble(
  key1 = setdiff(fct, after_other),
  key2= setdiff(after_other, fct)
  )

preserve_xmap <- tibble::tibble(
  key1 = intersect(fct, after_other),
  key2 = key1
)

xmap_set <- dplyr::bind_rows(preserve_xmap, collapse_xmap) |>
  dplyr::mutate(weight = 1)

identical(xmap_set, xmap_ref)
```

Visualising it

```{r code=readLines('R/plt_pm_df.R')}

```

```{r}
plt_pm_df(xmap_set, from = key1, to = key2, weights = weight)
```
